Individual heterogeneity in the functional topography of the DMN in medial prefrontal cortex 

### 2019-02-08 Manuscript.Rmd
first draft text ready, now working on figures

---
title: 'Individual heterogeneity in the functional topography of the DMN in medial prefrontal cortex '
author: "Claudio Toro-Serey, Sean Tobyne, & Joseph T. McGuire"
output:
  pdf_document: default
  word_document: 
    reference_docx: Style_reference.docx
  html_document: default
bibliography: mPFC_paper.bib
header-includes:
    - \usepackage{setspace}\doublespacing
---

@@

Where $I$ is the identity matrix, $D$ is a diagonal matrix containing the strength of each vertex (i.e. the sum of its correlation values with every other vertex), and $W$ is the adjusted correlation matrix. This resulted in a matrix wherein each entry was the normalized value of the connection (from 0 to 1) between any two vertices relative to their combined connectivity strength, and with ones along the diagonal. This transormation ensures that every row adds up to zero. We then computed the eigenvalues and eigenvectors of the normalized Laplacian, and used the eigenvector associated with the second-to-lowest eigenvalue to divide the network into two. This vector (from now on called the 'Fiedler vector') provides a set of positive and negative values to binarize the network with, and guarantees that the resulting communities are connected [@Fiedler1975]. Given that this data-driven method is agnostic to the prevalent function of each community (i.e. DMN or not DMN), we defined the DMN community to be the one containing the DMN-specific area from the meta-analysis (ensuring that the labeling of the DMN community was consistent across subjects). Importantly, given the high density of these networks (see results), spectral partitioning was unlikely to face the issues associated with its use in sparse networks [@Fortunato2016].

@@

The thresholded partitions in our work qualitatively resembled the topographical arrangement of networks A and B defined by Braga and Buckner [-@Braga2017]. We explored this similarity by reproducing their seed-based connectivity approach in three of our subjects. In their work, Braga and Buckner selected individual vertices along the dorsolateral prefrontal cortex (DLPFC) that produced these two anticorrelated networks. Therefore, we hypothesized that if the communities we identified on the medial wall match their proposed networks, we should be able to find seed vertices in DLPFC that can reproduce them. The whole-brain seed-based functional connectivity maps for these subjects were qualitatively compared to those produced by spectral partitioning.

@@

### *Correlation vs community detection in mPFC*

@@

Finally, we coarsely examined whether these idiosyncratic topographical arrangements were defined by anatomical brain features. To do so, we correlated each subject's Fiedler vector with their respective structural MRI data. On average, we found that the Fiedler vector is uncorrelated with myelin density (mean *r* = `r round(meanMyelin, digits = 2)`; SD = `r round(sdMyelin, digits = 2)`), cortical folding (mean *r* = `r round(meanCurv, digits = 2)`; SD = `r round(sdCurv, digits = 2)`), and surface thickness (mean *r* = `r round(meanThick, digits = 2)`; SD = `r round(sdThick, digits = 2)`). While these findings point at an orthogonality between structure and functional organization, they resonate with previous work that found no strong matches between intersubject functional variability and cortical thickness, and low variability in sulcal depth within DMN [@Mueller2013a].

@@

@@

```{r Thresholded maps (Fig 6), out.width = "80%", fig.align = "center", fig.cap= "Figure 6. A: Thresholded Fiedler vector maps for subject 100307 (top), and its binarized form (bottom). Values that did not meet the threshold were set to zero on the Fiedler vector, and 0.5 on the binarized maps (giving rise to a third community of high-variability vertices). B: Spatial mean of the binarized maps across all participants, indicating the proportion of DMN affiliations per vertex in our sample. This aggregate map shows the common organizational principle of the DMN and non-DMN communities, while also pinpointing the relatively high level of variability in mPFC."}
# Change to inflated eventually
include_graphics("./Images/thresholded.png")
```

@@

>>>>

---
title: 'Individual heterogeneity in the functional topography of the DMN in medial prefrontal cortex '
author: "Claudio Toro-Serey, Sean Tobyne, & Joseph T. McGuire"
output:
  word_document: 
    reference_docx: Style_reference.docx
  pdf_document: default
  html_document: default
bibliography: mPFC_paper.bib
header-includes:
    - \usepackage{setspace}\doublespacing
---

@@

Where $I$ is the identity matrix, $D$ is a diagonal matrix containing the strength of each vertex (i.e. the sum of its correlation values with every other vertex), and $W$ is the adjusted correlation matrix. This resulted in a matrix wherein each entry was the normalized value of the connection (from 0 to 1) between any two vertices relative to their combined connectivity strength, and with ones along the diagonal. This transormation ensures that every row adds up to zero. We then computed the eigenvalues and eigenvectors of the normalized Laplacian, and used the eigenvector associated with the second-to-lowest eigenvalue to divide the network into two. This vector (from now on called the 'Fiedler vector') provides a set of positive and negative values to binarize the network with, and guarantees that the resulting communities are connected [@Fiedler1975]. Every individual map is therefore binarized by sign in order to evaluate the topography of each sub-network (see next section for details). Figure 1 provides a schematic of the process. Given that this data-driven method is agnostic to the prevalent function of each community (i.e. DMN or not DMN), we defined the DMN community to be the one containing the DMN-specific area from the meta-analysis (ensuring that the labeling of the DMN community was consistent across subjects). Importantly, given the high density of these networks (see results), spectral partitioning was unlikely to face the issues associated with its use in sparse networks [@Fortunato2016].

@@

@@

We coarsely examined whether these idiosyncratic topographical arrangements were defined by anatomical brain features. To do so, we correlated each subject's Fiedler vector with their respective structural MRI data. On average, we found that the Fiedler vector is uncorrelated with myelin density (mean *r* = `r round(meanMyelin, digits = 2)`; SD = `r round(sdMyelin, digits = 2)`), cortical folding (mean *r* = `r round(meanCurv, digits = 2)`; SD = `r round(sdCurv, digits = 2)`), and surface thickness (mean *r* = `r round(meanThick, digits = 2)`; SD = `r round(sdThick, digits = 2)`). While these findings point at an orthogonality between structure and functional organization, they resonate with previous work that found no strong matches between intersubject functional variability and cortical thickness, and low variability in sulcal depth within DMN [@Mueller2013a].
### *Correlation vs community detection in mPFC*

@@

### *Similarities with other work*

The thresholded partitions in our work qualitatively resembled the topographical arrangement of networks A and B proposed by Braga and Buckner [-@Braga2017]. We explored this similarity by reproducing their seed-based connectivity approach in three of our subjects. In their work, Braga and Buckner selected individual vertices along the dorsolateral prefrontal cortex (DLPFC) that produced these two anticorrelated networks. Therefore, we hypothesized that if the communities we identified on the medial wall matched their proposed networks, we should be able to find seed vertices in DLPFC that can reproduce them. The whole-brain seed-based functional connectivity maps for these subjects are juxtaposed to their spectral partitioning counterparts on figure 15.

Visual inspection of these networks shows a remarkable similarity between our DMN and non-DMN communities and their networks A and B. These qualitative results are important, as here we reinforce the idea that canonical DMN can be reliably subdivide at the individual level. Further, it shows the potential of community detection methods to automatically extract these partitions without the need to rely on group data. One way in which our results are at odds with those from Braga and Buckner is the functional labeling of these sub-networks. Based on our criteria, network B would more closely identify with DMN (per 7m affiliation), while we currently do not have enough evidence to label network A. These discrepancies should be addressed with explicit behavioral manipulations in the future.

@@

```{r Method (Figure 1), out.width = "100%", fig.align = "center", fig.cap= "Summary of the partitioning scheme and its evaluation. Each connectome if transformed into a normalized Laplacian matrix, whose eigenvector associated with the second-to-lowest eigenvalue assigns negative or positive values to each vertex. These continuous maps are binarized by sign, and compared within and between subjects using the chance-correcting Adjusted Rand Index (ARI)."}
# Change to inflated eventually
include_graphics("./Images/method_figure.png")
```

@@

```{r Thresholded maps (Fig 6), out.width = "100%", fig.align = "center", fig.cap= "Figure 6. A: Thresholded Fiedler vector maps for subject 100307 (top), and its binarized form (bottom). Values that did not meet the threshold were set to zero on the Fiedler vector, and 0.5 on the binarized maps (giving rise to a third community of high-variability vertices). B: Spatial mean of the binarized maps across all participants, indicating the proportion of DMN affiliations per vertex in our sample. This aggregate map shows the common organizational principle of the DMN and non-DMN communities, while also pinpointing the relatively high level of variability in mPFC."}
# Change to inflated eventually
include_graphics("./Images/thresholded.png")
```

@@

```{r B&B comparison (Figure 15), out.width = "100%", fig.align = "center", fig.cap= "Qualitative comparison between maps produced by seed-based connectivity based on Braga & Buckner's (2017) criterion, and those produced by spectral partitioning. Networks A and B were produced by carefully selecting a DLPFC seed (pointed by the green arrows). In each case, our DMN and non-DMN networks map remarkably well to their proposed functional organization."}
# Change to inflated eventually
include_graphics("./Images/braga_comparison.png")
```
